Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-10 rebuttal.tex

\comment{page 1, column 1, line 24: the authors need to provide a
  reference for the assertion that cheminformatics is an older field,
  particularly given the comments later in this
  manuscript.}
\response{We have removed the assertion that cheminformatics is older
  than bioinformatics. Primarily, because there is no obvious
  reference that claims this. But alse because, depending on what one
  considers bioinformatics and cheminformatics, one field can be
  identified as being older or younger than the other. We feel that
  claims on age do not strengthen the paper and have thus removed it}

\comment{p1, c1, l29-37: I would dispute that until recently the
  cheminformatics techniques have been closely guarded
  secrets. J. Chem. Doc. was founded in 1961 and other techniques even
  older, such as the Wiener index (1947), Wiswesser Line Notation
  (1949), etc. Not to mention the actual foundations of what we call
  cheminformatics back to the atomistic theory of the 19th century.}
\response{We have the updated the text on page 1 to attenuate this
  statement, stressing on the fact that much of the data in
  cheminformatics is proprietary, whereas only some of the techniques (such
  as SMILES canonicalization) are. Regarding the statement that atomistic
  theory laid the foundations of cheminformatics - this is certainly
  true in a very broad sense. For that matter the atomistic theory
  laid the ground for computational chemistry in general. In the
  supplemental history document, we make a distinction between
  computational chemistry and cheminformatics}

\comment{p1, c1, l32: I think the phrase miracle molecule could be
  more usefully replaced with new drugs or new small molecule
  therapeutics.}
\response{The text has been updated to use the term ``therapeutic
  molecule''}

\comment{P3, c1, l41: the authors need to cover the Morgan algorithm
  (published in 1965) here and explain the canonicalization
  process. The issues mentioned in different canonicalization
  implementations providing different SMILES strings is also a
  challenge with InChI codes with different softwares giving different
  representation. Therefore, this is still not a solved problem as
  suggested here. Could the authors clarify this in the text?}
\response{We have updated the text to reference the Morgan
  algorithm. Regarding the issue of InChI codes, we believe that this
  is not the case. Currently, there is only one implementation of the
  InChI algorithm and even if alternative implementations were to be
  developed, the InChI specification is publically available. As a
  result, there should be no differing InChI representations for a
  given input structure.}

### 2011-11-10 paper.tex
Added more updates

While cheminformatics has a broad applicability, ranging from
agrochemical research and drug discovery to the design of novel
materials (since everything around us consists of ``molecules''), we have chosen to use drug discovery as the context of
this
article due to its central relevance for human wellbeing. Given that the safety of drugs on the market is a key concern
both for the general audience as well as researchers in pharmaceutical companies, the remainder of this article is
structured around the
concept of ``risk minimization'' in drug discovery---the question of how to minimize the chances of a small molecule
failing in the various research and development stages as a drug candidate, and how to successfully employ
cheminformatics methods in these efforts.  Table~\ref{table:properties} (in the Appendix) summarizes properties a drug candidate must satisfy to be considered ``promising,'' and sketches the role of cheminformatics at each stage of drug development.
%
\subsection{Databases \& Ontologies}
\label{sec:databases}
%
Most cheminformatics applications rely on large databases of chemical structures,
their properties, and relationships to, \textit{e.g.}, biological targets.
Organizing and maintaining such databases, as well as searching and clustering similar structures together,
are essential  to enable many scientific applications. Yet, each one of
those areas poses its unique computer science challenges.

>>>>

While cheminformatics has a broad applicability, ranging from
agrochemical research and drug discovery to the design of novel
materials, we have chosen to use drug discovery as the context of this
article due to its central relevance for human wellbeing. Given that
the safety of drugs on the market is a key concern both for the
general audience as well as researchers in pharmaceutical companies,
the remainder of this article is structured around the concept of
``risk minimization'' in drug discovery -- the question of how to
minimize the chances of a small molecule failing in the various
research and development stages as a drug candidate, and how to
successfully employ cheminformatics methods in these efforts.
Table~\ref{table:properties} (in the Appendix) summarizes properties a
drug candidate must satisfy to be considered therapeutically useful
(such as being appropriately soluble, exhibiting minimal toxicity and
being selective for the intended target) and sketches the role of
cheminformatics at each stage of drug development.
%
\subsection{Databases \& Ontologies}
\label{sec:databases}
%
Most cheminformatics applications rely on large databases of chemical structures,
their properties, and relationships to, \textit{e.g.}, biological targets.
Organizing and maintaining such databases, as well as searching and clustering similar structures together,
are essential  to enable many scientific applications. Yet, each one of
those areas poses its unique computer science challenges.
